More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-08-29 2016-Weak-FH-Bisim/WeakBisim.tex

\TODO{should we use iff/implies or LeftRightArrow/Rightarrow}
\begin{itemize}
\item[X] reorga sec weak bisim + thm -> 10 Juil
\item  intro et orga globale du papier -> fin aout
\end{itemize}

@@

\paragraph{Previous Works and Contribution}
While most of our previous works relied on closed, fully-instantiated semantics~\cite{BBCHM:article2009}, it is only recently that we could design a first version of a  parameterised semantics for pNets with a strong bisimulation equivalence~\cite{henrio:Forte2016}. This article builds open this previous parameterised semantics and provide a cleaner version of the semantics with a slightly simplified formalism and extends it to deal with weak bisimulation. The slight simplification of the formula allowed us to adopt a much cleaner presentation where the theory for open automata, the algebra used to define the semantics of pNets with holes, can be studied independently from its application to pNets. Also, in previous work the study of compositionability was only partial, and in particular the proof that bisimulation is an equivalence is one new contribution of this article and provides a particularly interesting insight on the semantic model we use.
The new formalism allowed us to extend the work and define weak bisimulation  for open automata, which is entirely new. This allows us to define a weak bisimulation equivalence for open pNets with valuable properties of compositionality. 
To summarise, the contribution of this paper are the following
\begin{itemize}
\item The definition of open automata: an algebra of parameterised automata with hole with a strong bisimulation equivalence. This is an adaptation of~\cite{henrio:Forte2016} with an additional property stating that strong bisimulation equivalence is indeed an equivalence relation.
\item A semantics for open pNets expressed as translation to open automata. This is an adaptation of~\cite{henrio:Forte2016} with a complete proof that strong bisimulation is compositional.
\item A theory of weak bisimulation for open automata, and its properties. It relies on the definition of weak open transitions that are derived from transitions of the open automaton by concatenating invisible action transitions with one (visible or not) transition. The precise definition of the concatenation is also a major contribution of this article.
\item A resulting weak bisimulation equivalence for open pNets and a simple static condition on synchronisation vectors inside pNets that is sufficient to ensure that weak bisimulation is compositional.
\item An illustrative example \TODO{TODO}
\end{itemize}

@@

@@

@@

@@

\begin{theorem}[FH-Bisimulation is an equivalence]\label{thm-equiv} Suppose $\mathcal{R}$ 
is an FH-bisimualtion. Then $\mathcal{R}$ is an equivalence, that is, $\mathcal{R}$ is 
reflexive, symmetric and transitive.

@@

       \section{Bisimulation is an equivalence: Proof of Theorem~\ref{thm-equiv}}
        \emph{Suppose $\mathcal{R}$ 
       	is an FH-bisimulation. Then $\mathcal{R}$ is an equivalence, that is, 
       	$\mathcal{R}$ is 
       	reflexive, symmetric and transitive.
       	}
%%% A PROOF WITH CONSTRUCTIVE AERGUMENTS BUT THAT DOES NOT WORK

>>>>

Global and questions:
\begin{itemize}
\item \TODO{should we use iff/implies or LeftRightArrow/Rightarrow}

\item \TODO{uniformiser pNet \pNet pNet s \pNet s (dans le texte et les formules!)}

\item \TODO{proposition Ludo \& Rabea: \pNet -> P,Q dans les formules pour que ce soit moins long}
\end{itemize}
Ludo:
\begin{itemize}
\item[X] reorga sec weak bisim + thm -> 10 Juil
\item  intro et orga globale du papier -> fin aout
\end{itemize}

@@

\paragraph{Previous Works and Contribution}
While most of our previous works relied on closed, fully-instantiated semantics~\cite{BBCHM:article2009,AmeurBoulifa2017,HKM-FASE16}, it is only recently that we could design a first version of a  parameterised semantics for pNets with a strong bisimulation equivalence~\cite{henrio:Forte2016}. This article builds open this previous parameterised semantics and provide a cleaner version of the semantics with a slightly simplified formalism and extends it to deal with weak bisimulation. The slight simplification of the formula allowed us to adopt a much cleaner presentation where the theory for open automata, the algebra used to define the semantics of pNets with holes, can be studied independently from its application to pNets. Also, in previous work the study of compositionability was only partial, and in particular the proof that bisimulation is an equivalence is one new contribution of this article and provides a particularly interesting insight on the semantic model we use.
The new formalism allowed us to extend the work and define weak bisimulation  for open automata, which is entirely new. This allows us to define a weak bisimulation equivalence for open pNets with valuable properties of compositionality. 
To summarise, the contribution of this paper are the following
\begin{itemize}
\item The definition of open automata: an algebra of parameterised automata with hole with a strong bisimulation equivalence. This is an adaptation of~\cite{henrio:Forte2016} with an additional property stating that strong bisimulation equivalence is indeed an equivalence relation.
\item A semantics for open pNets expressed as translation to open automata. This is an adaptation of~\cite{henrio:Forte2016} with a complete proof that strong bisimulation is compositional.
\item A theory of weak bisimulation for open automata, and its properties. It relies on the definition of weak open transitions that are derived from transitions of the open automaton by concatenating invisible action transitions with one (visible or not) transition. The precise definition of the concatenation is also a major contribution of this article.
\item A resulting weak bisimulation equivalence for open pNets and a simple static condition on synchronisation vectors inside pNets that is sufficient to ensure that weak bisimulation is compositional.
\item An illustrative example \TODO{TODO}
\end{itemize}

@@

\TODO{TODO ... or rather a dedicated section}

@@

pNets are composition entities equipped with a rich synchronisation mechanism: synchronisation vectors allows the expression of synchronisation between any number of entities and at the same time the passing of data between processes. Their strongest feature is that the data emitted by processes can be used can be used inside the synchronisation vector to do addressing: it is easy to synchronise a process indexed by $n$ with the action $a(v,n)$ of another process. This is very convenient to model systems and encode futures or message routing. pNets have been used to model GCM distributed component systems, illustrating the expressiveness of the model~\cite{AmeurBoulifa2017}. These works show that pNets are convenient to express the behaviour of the system in a compositional way, which is crucial for the definition of the semantics, especially when dealing with a hierarchical component system like GCM. Unfortunately, the semantics of pNets and the existing tools at this point were only able to deal with a closed system completely instantiated: pNets could be used as composition operator in the definition of the semantics, which was sufficient to perform finite-state model checking on a closed system, but there was no theory for the use of pNets as operators and no tool for proving properties on open system. The semmantics of closed pNets~\cite{AmeurBoulifa2017}, defined as an instantiation of labelled transition systems, is not necessary to understand this article but can be useful to  understand the semantics of pNets in a simpler and more operational setting. The theory of pNets as operators able to fully take into account open systems is given in the rest of this article.

@@

Though the definition is simple, the fact that transitions are complex structures relating events must not be underestimated in order to understand the rest of the article. The first element of theory for open automata, i.e. the definition of a strong bisimulation, is given below.

@@

\begin{theorem}[FH-Bisimulation is an equivalence]\label{thm-equiv} Suppose $\mathcal{R}$ 
is an FH-bisimualtion. Then $\mathcal{R}$ is an equivalence, that is, $\mathcal{R}$ is 
reflexive, symmetric and transitive.
\end{theorem}

The proof of this theorem can be found in Appendix~\ref{thm-equiv-proof}. The only non-trivial part of the proof is the proof of transitivity. It relies on the following elements. First,  the transitive composition of two relations with predicate is defined; this is not exactly standard as it necessitates to define the right predicate for the transitive composition and producing a single predicate to relate any two states. Then the fact that one open transition is simulated by a family of open transitions leads to a doubly indexed family of simulating open transition; this needs particular care, also because of the use of renaming (\Post) when proving that the predicates satisfy the definition (property on $\Pred_{s,t} \land \Pred_{OT}$ in the definition).

\TODO{have a look at the end of the sec}

@@

       \section{Bisimulation is an equivalence: Proof of Theorem~\ref{thm-equiv}}\label{thm-equiv-proof}
        \emph{Suppose $\mathcal{R}$ 
       	is an FH-bisimulation. Then $\mathcal{R}$ is an equivalence, that is, 
       	$\mathcal{R}$ is 
       	reflexive, symmetric and transitive.
       	}
%%% A PROOF WITH CONSTRUCTIVE AERGUMENTS BUT THAT DOES NOT WORK
